Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2014-07-22 writeups/dissertation/dissertation.tex
foo

\subsection{Aquiring Datasets}

@@

Twitter provides two types of APIs for aquiring data: A query-based and streaming interface.
The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
This question was tested by examing how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
The model predicts that the more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.

@@

For both equatons, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g,. the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.

The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occured most frequently will have the highest activation,
and activation depends less on the specific chunk that have been observed in the recent past.

@@

At each tagging instance, the model has activation values for a set of tags.
Model accurracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
Otherwise, it was marked as incorrect.

@@

Finally, model accuraccy was analyzed across all dataset slices for StackOverflow and Twitter.
Mean accuracy for both sites is included in Figure \ref{figPriorAcc}. 
The mean was computed by taking the mean of the mean model accuracy for each dataset slice.
All confidence intervals are 95\% bootstrapped for the mean of the means for each slice.

@@

However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
The constrained form of model accuracy can not be used for the regression because there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
The constrained form could be used in a regression if a custom form of the regression was implemented.
However this was not persued since around \num{2500} total regressions were run on these datasets for two reasons:
A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractible without a closed-form solution.

@@

The offset term ensures that the activation for the small number of top tags for each post are equal across posts.
This ensures that the logistic regression only labels a few 1s for each post.
\textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
However this method is complicated since computing the toal activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
\citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.

@@

However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
This dataset contains samples of a large set of various users that choose a specific set of hashtags being monitored.
So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
Since access rates are limited across that interface, it is infeasable to do this process for every user that appears in the Twitter popular-users dataset.
So instead, the global prior tag history for all users in the Twitter popular-users dataset was used to compute prior activation values for each tag.

@@

The context component for both the Random Permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
Already FIXME total co-occuring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrent matrix instead,
especially since FIXME co-occurrences are already used with \num{1000000} posts.

@@

In both plots, if the model name includes ``full'' then all model components are included.
When they are enumerated, then components are left out.
The random-permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
So, for example, the ``standard Prior model relaxed across posts w/o offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.

@@

It is actually somewhat dissapointing that the offset term is even needed.
If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to do for a specific post.
If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
However, since accuracy improves when adding the offset term for both models, it does not seem that activation values have much meaning across posts (and only within a single post) for these datasets.

A likely and simple explaination for this is that possibly there are posts where a large set of tags have high activation,
but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term actually happens.
It is unlikely that this retrieval threshold can be directly controlled and manipulated.
Rather, it could be a high-level task retrieval threshold that the author has control over, where he/she actively sets it based on competing goals of tagging posts an not overly saturating a post with tags.
Adding the offset term into the model is a way to implement a simple version of this high-level retrieval threshold.

@@

\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random-permutation model:
A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and uing a previously-derived stoplist hardly increased performance.
However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attentuating stop words based on Entropy was also explored.

@@

\paragraph{Determing Frequency Cutoff}

@@

The no-removal plot shows the size of the spikes when high-frequency words are removed.
The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
The cutoff was chosen such that the frequency plot looks similar to the entropy plot, with spikes around the same size.
Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
The cuttoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset were not given.

@@

The results show that model accuracy improves if stop words are handled in any of the three ways (pre-defined stoplist filtering, entropy weighting, and frequency filtering).
Using a pre-defined stoplist makes the smallest improvment, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.

@@

Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a pre-determined list.
However it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
However frequency filtering does have a few drawbacks compared to entropy weighting:
It produces accuracy measurments that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
Therefore, favoring parsimony, for the random-permutation model, attentuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.

@@

Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
This was not the case for the random-permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random-permutation model, and is most likely never the case for the random-permutation model.
Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random-permutation model does not have.
When computing activaiton for the random-permutation model, the corrollation is calculated directly on a matrix of counts of co-occurrences for words and tags.
However for the Bayesian model this count matrix is converted to a log odds ($S_{ji}$) matrix,
where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
So it is likely the case that a form of frequency \emph{attentuation} is already built into the Bayesian model.
And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.

\textcite{Stanley2013} found that the entropy-weighting technique to attentuate stop words for the Bayesian model improved model accuracy.
The same effect is found for the StackOverflow and Twitter datasets in this study.
Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
This was clearly the case for the Bayesian model.
Both frequency and entropy techniques performed equally for the random-permutation model.
However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random-permutation model for these datasets.

@@

The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
In order to see how accuracy changed as function of document size, 
co-occurence matrices for a range of smaller document sizes were built.
Both the random-permutation and Bayesian models were tested on all document sizes.

@@

Accuracy improves for both the Bayesian and random-permutation models as the size of the documents used to generate the co-occurrence matrix increases.
However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random-permutation model, while the frequency technique only begins to plateau. 
The Bayesian model does not plateau and accuracy continues to improveas document size increases.

@@

It is likely that the random-permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
Even if the non-predictive words based on entropy are attenuated,
when the high-frequency words are included random-permutatoin model,
this attentuation is not enough to overcome the range effects that grow with document size.
These effects end up saturating all of the correlations near one, and consequently decrease discriminability and accuracy.
However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
The Bayesian model does not saturate when using entropy weighting because each cell in the co-occurrence matrix is normalized by word frequency and tag frequency.

@@

Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateu for the random-permutation model with increasing document size.
Once again, the Bayesian model continues to improve with document size.

@@

It is interesting to compare model accuracy betwen the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
At smaller document sizes, the random-permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
In other words, the Bayesian model needs larger co-occurrence matrices than the random-permutation model to work properly.
As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random-permutation model is actually beneficial, and the compressed random-permutation model should be used. 
This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volitle, which happens at smaller document sizes.

@@

Three different row-size values were tested for the random-permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
A separate co-occurrence representation was built for each of these row values.
Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
Accuracy measurments were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.

@@

In the figure, the full random permutation model that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random-permutation model for all other results.
The plot is also broken out for different size datasets used to build the co-occurrence matrix.
Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
However \num{2048} rows is sufficient for these dataests, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.

@@

However it should be pointed out that the way the random-permutation model compresses the words is completly random.
It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
Everything is randomly chosen with the random-permutaton model.
The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
So it is that more impressive that even with a randomly-chosen form of compression, only \num{2048} rows are needed to represent all of the words in these datasets.

@@

The log-odds transformation technique was tested for the random-permutaiton model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
  \caption{Random-permutation model accuracy when using log-odds trasformation on context for StackOverflow}
  \label{figContextLogoddsSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
  \caption{Random-permutation model accuracy when using log-odds trasformation on context for Twitter}
  \label{figContextLogoddsT}
\end{figure}

@@

Vector-based models like the random-permutation model can easily incorporate word order when making predictions.
This is one of their primary strenghts that has been discussed \parencites{Jones2007, Sahlgren2008}.
In order to thouroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.

@@

Three different methods for incorporating word order in the random-permutation model were tested:
Taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
For the standard method, all words and their respective position are used to build the representation.
For the direction method, essentially two representations are combined: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
\textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
Teh same $+2$ $-2$ range was used to test the window method on the Twitter dataset.

Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for hte post.
The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.

@@

The original BEAGLE model \parencite{Jones2007} actually showed only minimal improvement when incorporating word order as well.
In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
\textcite{Sahlgren2008}'s results for word order for the random-permutation model are similar.
At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model, which does not make sense.
\textcite{Sahlgren2008} most likely did not use a regression technique to opimize the weights for each of the order and orderless model terms.
So although incorporating word order into the random-permutation model does seem to slightly improve performance,
the strenght of the model is more likely to be in how effeciently it can represent the co-occurrence matrix, than its ability to include word order in the representation.

Since it was shown that including word-order information has little to no sigificant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.

@@

For both the Bayesian and random-permutation models, accuracy remains the same when the prior compoennt is ermoved for StackOverflow.
Accuracy does dramatically decrease for Twitter when the prior is removed, however.

@@

This shows that incorporating past tagging history is important for even a site like StackOverflow where a large amount of context information has been collected.
However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predicitve variance.

>>>>

\subsection{Acquiring Datasets}

@@

Twitter provides two types of APIs for acquiring data: A query-based and streaming interface.
The \textcite{Tweepy} Python package provides a Python functional interface to these APIs, and this package was used to collect data from Twitter.
The query-based interface was used to collect all tweets for specific popular users.
At most \num{3000} of the most recent tweets were collected for each user, since Twitter only allows access to this number of tweets for each user.

@@

How does a user's prior hashtag use influence their likelihood of choosing that hashtag in the future?
This question was tested by examining how well ACT-R's base-level activation component of declarative memory predicts the tags that authors choose, given each author's past tagging behavior.
The model predicts that the more recently and frequently used chunks have higher prior activations, and are more likely to be needed in the future.

@@

For both equations, $i$ represents each chunk (e.g., a word), $j$ is a specific observation for a chunk (e.g., the 2nd time that ``a'' appears), and $n$ is the number of times each chunk has been observed.
Since the optimized learning form assumes a constant presentation rate for each chunk,
it only has to keep track of the time since the first presentation of a chunk $L$ and the total number of instances observed for that chunk. 
It does not have to keep track of the time each specific chunk instance was observed.
Consequently, the optimized learning form of $B_{i}$ is computationally much less expensive than the standard form, and is actually set to the default in the ACT-R common-lisp implementation.

The decay rate parameter $d$ represents how fast the activation for an individual presentation of a chunk decays from declarative memory.
Higher values mean that chunk activation decays faster, which means that the most recently-presented chunks will have the highest activation, and the total number of times that a chunk has been presented matters less.
Lower decay rate values flip the importance of frequency and recency: Lower values mean that chunks that have occurred most frequently will have the highest activation,
and activation depends less on the specific chunk that have been observed in the recent past.

@@

At each tagging instance, the model has activation values for a set of tags.
Model accuracy was assessed by rank ordering these activations and choosing the model's highest $N$ tags, where $N$ is the number of tags used for that specific post.
The model's chosen $N$ tags were compared to the user's chosen $N$ tags, and if a model's chosen tag was in the set of user's chosen tags, then the prediction was marked as correct.
Otherwise, it was marked as incorrect.

@@

Finally, model accuracy was analyzed across all dataset slices for StackOverflow and Twitter.
Mean accuracy for both sites is included in Figure \ref{figPriorAcc}. 
The mean was computed by taking the mean of the mean model accuracy for each dataset slice.
All confidence intervals are 95\% bootstrapped for the mean of the means for each slice.

@@

However, this relaxed metric was used for the rest of the experiments that involve combining terms because the optimal weights found by the regression are based on using the relaxed metric.
If these weights are then used to compute the constrained form of model accuracy, biases may be introduced since the regression wasn't optimized to the constrained form, and weights may change if that was done.
The constrained form of model accuracy can not be used for the regression because there is no current logistic regression technique in R that can be constrained to label a specific number of 1s for each observation. 
The constrained form could be used in a regression if a custom form of the regression was implemented.
However this was not pursued since around \num{2500} total regressions were run on these datasets for two reasons:
A custom implementation in R would not likely run in compiled C, and even if it was implemented in a compiled language it would be computationally intractable without a closed-form solution.

@@

The offset term ensures that the activation for the small number of top tags for each post are equal across posts.
This ensures that the logistic regression only labels a few 1s for each post.
\textcite{Stanley2013} did this by subtracting off the mean total activation for the top 5 tags for each post.
However this method is complicated since computing the total activation for a tag requires already knowing the weights for each model component that are used to compute that activation.
\citeauthor{Stanley2013} used an iterative approach to determine the weights used for the offset term, and kept iterating until those weights matched the weights returned after running the logistic regression.

@@

However, a custom user prior for the Twitter popular-hashtags dataset cannot be used, since the entire Twitter dataset cannot be downloaded.
This dataset contains samples of a large set of various users that choose a specific set of hashtags being monitored.
So in order to generate a custom user prior for each sample in the dataset, that user's entire tagging history must be accessed through the API.
Since access rates are limited across that interface, it is infeasible to do this process for every user that appears in the Twitter popular-users dataset.
So instead, the global prior tag history for all users in the Twitter popular-users dataset was used to compute prior activation values for each tag.

@@

The context component for both the Random Permutation and Bayesian model requires a co-occurrence matrix of words and observed tags.
Different size matrices were analyzed, ranging from \num{1000}, \num{10000}, \num{100000}, \num{1000000}, and \num{3000000} posts.
The matrix for \num{3000000} posts was only used for Twitter, since the number of words in a StackOverflow post is much higher than Twitter on average,
and the space and computational time required when using a co-occurrence matrix of \num{3000000} posts for StackOverflow is too high for a current high-performance desktop, even with highly optimized representations.
Already FIXME total co-occurring word tag observations are used to build a co-occurrence matrix of \num{1000000} StackOverflow posts.
It is also the case that \num{1000000} posts for StackOverflow is already around $1/7$ of the total posts created on the site,
so it is reasonable to assume that the results would not change greatly for StackOverflow if $3/7$ of the total posts created were used to build the co-occurrence matrix instead,
especially since FIXME co-occurrences are already used with \num{1000000} posts.

@@

In both plots, if the model name includes ``full'' then all model components are included.
When they are enumerated, then components are left out.
The random-permutation models are represented as ``RP'', Bayesian models as ``Bayes'', and they are plotted alongside each other so that comparisons across the two model types can be done easily.
So, for example, the ``standard Prior model relaxed across posts without offset'' means that only the prior model component was used, and the offset for that component was not subtracted off when determining activation.
As another example, the ``Bayes only orderless context'' means that only the context component of the Bayesian model was used (along with the offset), and that no order information was used for that contextual component term.

@@

It is actually somewhat disappointing that the offset term is even needed.
If it were the case that model activation for a tag in a specific post could be interpreted as an exact activation value that did not depend on the specific post,
then the model would be able to say when, for example, the third top tag in a post is more likely to be a chosen tag than the second top tag in another post.
That is, the activation values would have meaning more than just rank ordering, and could be used to determine, for example, the number of retrievals to do for a specific post.
If this were the case, then adding the offset term would actually decrease accuracy, since this absolute value of the activation values would be subtracted off.
However, since accuracy improves when adding the offset term for both models, it does not seem that activation values have much meaning across posts (and only within a single post) for these datasets.

A likely and simple explanation for this is that possibly there are posts where a large set of tags have high activation,
but the author is throttling their retrieval threshold for each post such that only a small set of tags are actually used.
This is probably not the retrieval threshold used in the core declarative memory system that determines if a retrieval for the term actually happens.
It is unlikely that this retrieval threshold can be directly controlled and manipulated.
Rather, it could be a high-level task retrieval threshold that the author has control over, where he/she actively sets it based on competing goals of tagging posts an not overly saturating a post with tags.
Adding the offset term into the model is a way to implement a simple version of this high-level retrieval threshold.

@@

\textcite{Sahlgren2008} looked at several ways to handle the commonly-occurring stop words for the random-permutation model:
A data-driven frequency-weighted approach was tried, along with using a previously-derived stoplist, and no stop word removal at all.
Removing the high-frequency words in the dataset showed the highest performance increase compared to no stop word removal, and using a previously-derived stoplist hardly increased performance.
However, \citeauthor{Sahlgren2008} did not try to weight the stop words, and only tried methods to remove them.
So alongside looking at frequency weighting, previously derived, and no stop word removal, \textcite{Stanley2013}'s method of attenuating stop words based on Entropy was also explored.

@@

\paragraph{Determining Frequency Cutoff}

@@

The no-removal plot shows the size of the spikes when high-frequency words are removed.
The entropy plot shows how the size decreases by more than an order of magnitude when the entropy weighting measure is used.
The cutoff was chosen such that the frequency plot looks similar to the entropy plot, with spikes around the same size.
Note that using a previously-defined stoplist does not remove some of the remaining high-frequency words in these datasets, which hints that this method may not be ideal when model accuracy is examined.
The cutoff that produced the frequency plot is where a word represents more than \num{.0004} occurrences in the dataset.
This may or may not be equal to the value for \textcite{Sahlgren2008}, since that number was expressed in total observations (not as a percentage), and the total number of words in the dataset were not given.

@@

The results show that model accuracy improves if stop words are handled in any of the three ways (pre-defined stoplist filtering, entropy weighting, and frequency filtering).
Using a pre-defined stoplist makes the smallest improvement, while the entropy and frequency techniques improve performance the most, and frequency shows a slight edge over entropy.
However, using the frequency-weighting technique produces more noise in accuracy measurements compared to the other techniques, at least when tested on words in the title of StackOverflow posts.

@@

Comparing across Figures \ref{figContextStopRPSO} and \ref{figContextStopRPT}, entropy and frequency techniques are superior to filtering based on a pre-determined list.
However it is more difficult to choose the best approach when comparing entropy weighting and frequency filtering.
Frequency filtering is slightly better for StackOverflow, while entropy weighting is slightly better for Twitter.
However frequency filtering does have a few drawbacks compared to entropy weighting:
It produces accuracy measurements that have higher levels of noise (see Figure \ref{figContextStopRPSO} and frequency models for the title component in StackOverflow).
Also, the ideal cutoff threshold must be identified in order to filter based on frequency,
and identifying this cutoff requires a search, and this search must be done for each dataset that frequency filtering is used on.
This threshold is a free parameter in frequency filtering that is not required when weighting stop words by entropy.
Therefore, favoring parsimony, for the random-permutation model, attenuating stop words based on entropy produces more accurate results when compared to filtering the stop words based on frequency of use.
This result is interesting since filtering stop words based on frequency is the default method for random permutation models \parencite{Sahlgren2008}.
This suggests that a simpler method (entropy weighting) can be used instead of the more common frequency-filtering technique to achieve the same (or better) accuracy measurements.

@@

Another interesting result was how the Bayesian model actually performed worse when stop words were removed based on frequency.
This was not the case for the random-permutation model, was not the case when \textcite{Sahlgren2008} used frequency filtering on the random-permutation model, and is most likely never the case for the random-permutation model.
Model accuracy most likely decreases for the Bayesian model because this model already has a normalization component built in that the random-permutation model does not have.
When computing activation for the random-permutation model, the correlation is calculated directly on a matrix of counts of co-occurrences for words and tags.
However for the Bayesian model this count matrix is converted to a log odds ($S_{ji}$) matrix,
where both the number of observations for the word and the tag are normalized when computing the value for each cell (see the $S_{ji}$ computation in Table \ref{tabModACTRModel}).
So it is likely the case that a form of frequency \emph{attenuation} is already built into the Bayesian model.
And therefore adding a frequency filter on top of this does nothing to improve accuracy, and only removes predictive information.

\textcite{Stanley2013} found that the entropy-weighting technique to attenuate stop words for the Bayesian model improved model accuracy.
The same effect is found for the StackOverflow and Twitter datasets in this study.
Further, when this entropy-weighting technique is compared directly to the more common frequency-filtering technique, weighting by entropy produces more accurate results.
This was clearly the case for the Bayesian model.
Both frequency and entropy techniques performed equally for the random-permutation model.
However since the entropy-weighting technique is parameter free and does not require a cutoff to tune, it is also a better choice for the random-permutation model for these datasets.

@@

The default number of documents used to build the co-occurrence matrix was \num{1000000} for StackOverflow and \num{3000000} for Twitter.
In order to see how accuracy changed as function of document size, 
co-occurrence matrices for a range of smaller document sizes were built.
Both the random-permutation and Bayesian models were tested on all document sizes.

@@

Accuracy improves for both the Bayesian and random-permutation models as the size of the documents used to generate the co-occurrence matrix increases.
However model accuracy using the entropy weighting technique plateaus and slightly decreases for the largest document size for the random-permutation model, while the frequency technique only begins to plateau. 
The Bayesian model does not plateau and accuracy continues to improves document size increases.

@@

It is likely that the random-permutation model is saturating when entropy weighting is used for the largest document size for StackOverflow.
Even if the non-predictive words based on entropy are attenuated,
when the high-frequency words are included random-permutation model,
this attenuation is not enough to overcome the range effects that grow with document size.
These effects end up saturating all of the correlations near one, and consequently decrease discrimination and accuracy.
However, the effect of saturation is small compared to the large effect from using either frequency filtering or entropy weighting to handle the stop words.
The Bayesian model does not saturate when using entropy weighting because each cell in the co-occurrence matrix is normalized by word frequency and tag frequency.

@@

Both entropy weighting and frequency filtering in Figure \ref{figContextDSizeSO} plateau for the random-permutation model with increasing document size.
Once again, the Bayesian model continues to improve with document size.

@@

It is interesting to compare model accuracy between the Bayesian and random-permutation model at different document sizes for both StackOverflow and Twitter.
At smaller document sizes, the random-permutation model outperforms the Bayesian model (slightly for Twitter, dramatically for StackOverflow), and this performance difference deteriorates as document size increases.
In other words, the Bayesian model needs larger co-occurrence matrices than the random-permutation model to work properly.
As the size of the dataset increases, the compression may start to lose crucial information, and the uncompressed Bayesian model should be used.
As the size of the dataset decreases, the noise and crosstalk introduced by the compression for the random-permutation model is actually beneficial, and the compressed random-permutation model should be used. 
This crosstalk and noise is beneficial because it helps soften spurious $S_{ji}$ values when the values for each individual cell in the co-occurrence matrix are still volatile, which happens at smaller document sizes.

@@

Three different row-size values were tested for the random-permutation model to see how accuracy changed as a function of space required to represent the co-occurrence matrix.
A low value (\num{100}), a more standard value (\num{2048}), and a high value (\num{10000}) of rows were tested.
A separate co-occurrence representation was built for each of these row values.
Each representation was tested on the four Twitter popular-hashtags datasets and the StackOverflow randomly-sampled dataset.
Accuracy measurements were the same as were used for previous runs of these datasets: mean of the mean accuracy for each of the 5 StackOverflow and 15 Twitter runs.

@@

In the figure, the full random permutation model that does not have the number of rows in the matrix specified is the \num{2048}-row matrix.
This was done so that names in all of the plots would remain consistent across all plots, and the \num{2048}-row matrix is the size used for the random-permutation model for all other results.
The plot is also broken out for different size datasets used to build the co-occurrence matrix.
Reducing compression to only 100 rows in the representation has a strong negative effect on accuracy.
However \num{2048} rows is sufficient for these datasets, as increasing the number of rows to almost an order of magnitude higher (\num{10000}) has little to no positive impact on accuracy.

@@

However it should be pointed out that the way the random-permutation model compresses the words is completely random.
It is not an LSA approach where the few ones and negative ones for each word are chosen such to maximize the ability for the compressed matrix to represent the full matrix.
Everything is randomly chosen with the random-permutation model.
The environment vector for each word is randomly selected, and no SVD is performed to choose these in a more discriminating manner.
So it is that more impressive that even with a randomly-chosen form of compression, only \num{2048} rows are needed to represent all of the words in these datasets.

@@

The log-odds transformation technique was tested for the random-permutation model on the Twitter popular-hashtags and StackOverflow randomly-sampled datasets.

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
  \caption{Random-permutation model accuracy when using log-odds transformation on context for StackOverflow}
  \label{figContextLogoddsSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-logoddsSO-crop.pdf}}}
  \caption{Random-permutation model accuracy when using log-odds transformation on context for Twitter}
  \label{figContextLogoddsT}
\end{figure}

@@

Vector-based models like the random-permutation model can easily incorporate word order when making predictions.
This is one of their primary strengths that has been discussed \parencites{Jones2007, Sahlgren2008}.
In order to thoroughly test the impact of adding word order into the model, various methods for incorporating word order were tested on the StackOverflow and Twitter datasets.

@@

Three different methods for incorporating word order in the random-permutation model were tested:
Taking just the sign of the word's position relative to the tag (direction method), using the position (standard method), and using the position but only including words that were used near the tag (window method).
For the standard method, all words and their respective position are used to build the representation.
For the direction method, essentially two representations are combined: all co-occurrences with words to the left of the tag, and all co-occurrences with words to the right of the tag.
The window method is similar to the standard method, where each word's relative position is maintained, but it only includes words near the tag to build the representation.
\textcite{Sahlgren2008} tested similar methods, and found that accuracy was highest when a narrow window was used for the window method (only including words at most two positions to the left and right of the tag).
The same $+2$ $-2$ range was used to test the window method on the Twitter dataset.

Only the Twitter dataset was tested, since the StackOverflow dataset contains no word order information between words in the title and body of the post and the tags used for the post.
The author chooses tags after filling out the title and body forms in the post, so the tags are never intermixed with words in the post like they are for Twitter.

@@

The original BEAGLE model \parencite{Jones2007} actually showed only minimal improvement when incorporating word order as well.
In absolute terms, this model only improved by 2 percentage points in accuracy when trained on the TASA and tested on the TOEFL.
\textcite{Sahlgren2008}'s results for word order for the random-permutation model are similar.
At a row size of around \num{2000}, model accuracy improves by less than a percentage point when order information is incorporated.
The results for larger row sizes are somewhat inconsistent, since the full model (order and orderless) sometimes underperforms the order-only model, which does not make sense.
\textcite{Sahlgren2008} most likely did not use a regression technique to optimize the weights for each of the order and orderless model terms.
So although incorporating word order into the random-permutation model does seem to slightly improve performance,
the strength of the model is more likely to be in how efficiently it can represent the co-occurrence matrix, than its ability to include word order in the representation.

Since it was shown that including word-order information has little to no significant improvement on model accuracy, an attempt to incorporate word order information into the Bayesian model was not made.

@@

For both the Bayesian and random-permutation models, accuracy remains the same when the prior component is removed for StackOverflow.
Accuracy does dramatically decrease for Twitter when the prior is removed, however.

@@

This shows that incorporating past tagging history is important for even a site like StackOverflow where a large amount of context information has been collected.
However, for this type of site it only becomes important when a reasonably large amount of a user's past tagging history has been collected.
For a randomly-sampled user on StackOverflow, a context-only model captures the majority of the predictive variance.
